A thesis draft

### 2018-11-11 res/sections/08-Background.tex
Fix SDN stack subsection

\subsection{The SDN stack}
\begin{figure}[t]
 \centering
 \includegraphics[scale=1.3]{sdn_architecture}
 \caption{SDN architecture}
 \label{chap:background:img:sdn_architecture}
\end{figure}
As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
architecture can be splitted in three different layers, based on their
functionalities. The \emph{SDN Applications} role is to set the 
behaviour of the SDN controllers and manage its hardware resources and data 
through APIs, often called \emph{northbound interfaces}(NBI).
These interfaces provide an additional abstraction layer that high-level 
applications can consume, enabling the possibility to use a variety of 
different verdors' controllers in the same way.
Analysing the SDN applications is possible to note two defined layers, one that 
is responsible to manage the whole logic, and another one dedicated to the 
communication part, that in particular uses the previsously defined NBIs.
Theese applications can also build an abstracted view of the network, using
controllers information, in order to make the traffic analysis easier and to 
simplify the decision making process.

\emph{SDN controllers}, instead, are logically centralized components that
receive requests from the application and either translate them in a fashion
that can be understood from the lower layer, or collect data needed by the
application. Because of they have to accomplish theese aims, they can be see
as 3 layers programs, in which a layers exposes an interface to the upper
applications (\emph{NBI agents}), the second one represent the SDN controller
logic and finally the third layer has the task to communicate with the network
(Control to Data-Plane Interface or southbound interface).

In the lowest part of the architecture we can find the \emph{SDN Networking 
Devices}. These ones control packet forwarding and data processing capabilities 
of the network. They comprises a CDPI agent and a set of one or more traffic 
forwarding engines and zero or more traffic processing functions (the logic of 
the devices).

\subsection{The OpenFlow protocol}
The SDN strategy is often associated with OpenFlow protocol: this communication
protocol aim is to give access to the forwarding network component to the lower
layer network infrastructure, made of switches and routers. OpenFlow enables
the remote administration of networks working on layer 3: packets forwarding
tables can be modified relying on packet matching rules. Following this
approach it is possible to change routes periodically or ad-hoc, based on
packet types, traffic or network load.

>>>>

\subsection{SDN stack}
\begin{figure}[t]
 \centering
 \includegraphics[scale=1.3]{sdn_architecture}
 \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
denote the three layer logic: the Application plane, the Controller plane and 
the Data plane.}
 \label{chap:background:img:sdn_architecture}
\end{figure}
As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
architecture can be splitted in three different layers, called planes, based on 
their functionalities.

\subsubsection{Application plane}

The \emph{Application plane} role is to set the behaviour of the SDN controllers 
and manage its hardware resources and data through APIs, often called 
\emph{northbound interfaces} (NBI).
These interfaces provide an additional abstraction layer that high-level 
applications can consume, enabling the possibility to use a variety of 
different verdors' controllers in the same way.
Analysing the SDN applications is possible to note two defined layers, one that 
is responsible to manage the whole logic, and another one dedicated to the 
communication part, that in particular uses the previsously defined NBIs.
These applications can also build an abstracted view of the network, using
controllers information, in order to make the traffic analysis easier and to 
simplify the decision making process.

\subsubsection{Controller plane}

\emph{SDN controllers}, are logically centralized components that
receive requests from the mentioned SDN applications and either translate them 
in a fashion that can be understood from the lower layer\todo{who?}, or collect 
data needed by the application itself. 
Due to the variety of tasks a controller has to accomplish, its logical 
suddivision is achieved with a three-layer architecture, in which every layer 
exposes it's own interface, that can be consequently used by the upper 
layers, up to the \emph{NBI agents}\todo{explain them?}.
In particular, the middle layer represents the controller inner-logic, which 
process the requests coming from the upper layer (aka the third layer), that 
has the specific task to communicate with the remaining network compoents, in 
other words the Control to Data-Plane Interface and/or the already 
mentioned southbound interface.

\subsubsection{Data plane}

In the lowest part of the architecture we can find the \emph{SDN Networking 
Devices}. These ones controls the packet forwarding and data processing 
capabilities of the network. They consist of a CDPI agent and a set of one or 
multiple traffic forwarding engines and of a zero or more traffic processing 
functions, that are, as a matter of fact, the devices logic.

\subsection{OpenFlow protocol}
The SDN strategy is often associated with OpenFlow protocol: this communication
protocol aims is to give access to the forwarding network component to the lower
layer network infrastructure, made of switches and routers. OpenFlow enables
the remote administration of networks working on layer 3: packets forwarding
tables can be modified relying on packet matching rules. Following this
approach it is possible to change routes periodically or ad-hoc, based on
packet types, traffic or network load.
